MCC Visualizations

### 2019-01-09 stats/adults/ex3_comparisons.Rmd
first draft CogSci paper

load("mcc_model_comparison.RData")

@@

@@

```{r model loglikelihoods}
###################
# RSA Models
###################

@@

@@

@@

@@

```{r}
save.image("mcc_model_comparison.RData")
```

>>>>

load("../saves/mcc_model_comparison.RData")

@@

```{r model no prior noise loglikelihood}
no.prior.loglikelihood <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.allPriorProbs

@@

```{r model loglikelihoods}

###################
# RSA Models
###################

@@

###################
# no Prior Models
###################

llh.no.prior <- webppl(
  program_code = paste(rsaUtils, rsaModel2, no.prior.loglikelihood, sep='\n'),
  data = list(allData = data, allConditions = allConditions, allPriorProbs = allPriorProbs), 
  data_var = "dataFromR",
  model_var = "model",
  inference_opts = list(method = "forward", samples = 10000),
  chains = 3, cores = 3
)%>%
  mutate(Model = "No Prior")

@@

```{r bayes factors, echo = FALSE}

bf <- data.frame(
  Comparison = c("Pragmatic vs. Prior only","Pragmatic vs. No prior", "No Prior vs. Prior only"),
  BayesFactor = c(
    exp(llh.plot$logP[llh.plot$Model== "Pragmatic"] - llh.plot$logP[llh.plot$Model== "Prior only"]),
    exp(llh.plot$logP[llh.plot$Model== "Pragmatic"] - llh.plot$logP[llh.plot$Model== "No Prior"]),
    exp(llh.plot$logP[llh.plot$Model== "Prior only"] - llh.plot$logP[llh.plot$Model== "No Prior"]))
)

bf%>%
  kable(digits = 2)
```

@@

```{r}
save.image("../saves/mcc_model_comparison.RData")

saveRDS(bf, "../saves/ex3_bf.rds")

saveRDS(vp.bf, "../saves/ex4_bf.rds")

# old code .. ignore

@@

### 2019-01-09 stats/adults/mcc_priors_overview.Rmd
first draft CogSci paper

```{r data}
d3 <- read_csv(file="novel.weak.data2.csv") %>%
  mutate(trial_type = ifelse(trial == "train", "train", "test"),
         strength = "2_medium")

d4 <- read_csv(file="novel.weak.data.csv") %>%
  mutate(trial_type = ifelse(trial == "train", "train", "test"),
         strength = "1_weak")

d1 <- read_csv(file="novel.data.csv") %>%
  mutate(trial_type = ifelse(trial == "train", "train", "test"),
         strength = "3_strong")

d2 <-read_csv(file="pref.data.csv") %>%
  mutate(trial_type = ifelse(trial == "train", "train", "test"),
         strength = "3_strong")

d5 <-read_csv(file="pref.weak.data.csv") %>%
  mutate(trial_type = ifelse(trial == "train", "train", "test"),
         strength = "2_medium")

@@

d <- d %>% 
  distinct(id, alltrial, .keep_all = TRUE)
```

```{r plot data}

>>>>

```{r data, include = F}
d3 <- read_csv(file="../data/novel.weak.data2.csv") %>%
  mutate(trial_type = ifelse(trial == "train", "train", "test"),
         strength = "2_medium")

d4 <- read_csv(file="../data/novel.weak.data.csv") %>%
  mutate(trial_type = ifelse(trial == "train", "train", "test"),
         strength = "1_weak")

d1 <- read_csv(file="../data/novel.data.csv") %>%
  mutate(trial_type = ifelse(trial == "train", "train", "test"),
         strength = "3_strong")

d2 <-read_csv(file="../data/pref.data.csv") %>%
  mutate(trial_type = ifelse(trial == "train", "train", "test"),
         strength = "3_strong")

d5 <-read_csv(file="../data/pref.weak.data.csv") %>%
  mutate(trial_type = ifelse(trial == "train", "train", "test"),
         strength = "2_medium")

@@

d <- d %>% 
  distinct(id, alltrial, .keep_all = TRUE)

length(unique(d$id))
```

```{r plot data, echo = F}

### 2019-01-09 stats/adults/visualizations.Rmd
first draft CogSci paper

```{r setup, include=FALSE}
load("mcc_model_comparison.RData")

@@

## pointrange plot
p1 <- d %>%
  filter(trial_type == "test") %>%
  mutate(control = ifelse(control == "true", "Control", "Test")) %>%
  group_by(condition ,control, id) %>%
  summarise(correct = mean(correct))

p2 <- p1 %>%
  multi_boot_standard(col = "correct") %>%
  ungroup()%>%
  mutate(condition = as.character(condition))

@@

exp12 <- ggplot() +
  geom_rect(data = p2, aes(fill = NA), alpha = 0.4 ,xmin = -Inf,xmax = Inf, ymin = -Inf,ymax = Inf)+
  geom_jitter(data = p1, aes(x = control, y = correct, col = control), alpha = 0.2, width = .3,height = .025)+
  geom_pointrange(data = p2, aes(x = control, y = mean, col = control,ymin = ci_lower, ymax = ci_upper),size = 1)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Expected Choice")+
  facet_wrap( ~ condition, labeller = as_labeller(c(`2`="Exp. 2 - Preference", `3`="Exp. 2 - Novelty", `1`="Exp. 1 - Informativeness"))) +
  theme_few(base_size = 20) + 
  ylim(-0.05,1.05)+
  guides(alpha = F, fill = F)+ 
  scale_color_manual(name="Condition",
                     breaks=c("Control","Test"),
                     values= c("#657b83","black"))+
  scale_fill_manual(name="Condition",
                     breaks=c("1","2","3"),
                     values= c("#dc322f","#268bd2","#268bd2"))+
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank(),
        panel.border = element_rect(color = "#268bd2", fill = NA, size = 6))

@@

@@

>>>>

```{r setup, include=FALSE}
load("../saves/mcc_model_comparison.RData")

@@

## pointrange plot
p1 <- d %>%
  filter(trial_type == "test") %>%
  mutate(control = ifelse(control == "true", "Control", "Test"),
         condition = factor(condition), 
         experiment = ifelse(condition == "1", "1","2"))%>%
  group_by(experiment,condition ,control, id) %>%
  summarise(correct = mean(correct))

@@

 ggplot() +
  geom_rect(data = p2, aes(colour = experiment, size = 4), fill = NA, alpha = 1 ,xmin = -Inf,xmax = Inf, ymin = -Inf,ymax = Inf, show.legend=FALSE)+
  geom_jitter(data = p1, aes(x = control, y = correct, col = control), alpha = 0.2, width = .3,height = .025)+
  geom_pointrange(data = p2, aes(x = control, y = mean, col = control,ymin = ci_lower, ymax = ci_upper),size = 1)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="",y="Expected Choice")+
  facet_wrap( ~ condition, labeller = as_labeller(c(`2`="Exp. 2 - Preference", `3`="Exp. 2 - Novelty", `1`="Exp. 1 - Informativeness")))+
  theme_few() + 
  ylim(-0.05,1.05)+
  guides(alpha = F,size =F)+
  scale_color_manual(name = "Condition",
                     breaks=c("Test","Control"),
                     values= c("#dc322f","#268bd2","#839496","black"))+
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank(),
        panel.border = element_blank()
        )

@@

```{r}
ggarrange(exp1,exp2,exp3, ncol = 3, nrow = 1, legend = "right", common.legend = T)
```

@@

# kids

```{r pretest data, include=FALSE}
pref_data <- read_csv(file="../data/kids_pref_data.csv")%>%
  filter(age != "2",
         trial_type != "train")%>%
  mutate(age_bin = factor(age),
         experiment = "preference_pretest",
         condition = ifelse(change == "false", "same_speaker", "different_speaker"),
         agent = ifelse( change == "false", agent, altAgent),
         minage = min(age_num))%>%
  select(experiment,subid,age_bin,age_num,trial,condition,agent,correct,minage)

inf_data <- read_csv(file="../data/kids_inf_data.csv")%>%
  filter(subage != "2",
         trial != "filler1",
         trial != "filler2")%>%
  mutate(age_bin = factor(subage),
         experiment = "informativeness_pretest",
         condition = ifelse(control == "no", "test", "control"),
        minage = min(age_num))%>%
  select(experiment,subid,age_bin,age_num,trial,condition,agent,correct,minage)

pre_data <- bind_rows(pref_data,
                      inf_data)
```

```{r pretest plot so, echo = F}

pre_so <- pre_data %>%
  filter(experiment == "informativeness_pretest")%>%
  group_by(age_num,subid)%>%
  summarise(correct = mean(correct))

ggplot(data = pre_so, aes(x = age_num, y = correct)) +
  geom_jitter(alpha = 1, height = 0.1, width = 0.1, col = "grey")+
  geom_smooth(method = "lm", se = T, col = "black",fill = "darkgreen", alpha = .4, size = 1)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="age",y="Proportion Expected Choice")+
  theme_few(base_size = 20) +
  #ggtitle("Informativeness data")+
  ylim(-0.05,1.05)+
  guides(alpha = F)+ 
  coord_fixed(ratio = 2)+
  theme(panel.border = element_rect(color = "#dc322f", fill = NA, size = 6))
```

```{r}
ggsave("../graphs/kids_inf2.png", width = 3, height = 3, scale = 1.5)
```

```{r pretest plot prior, echo = F}

pre_prior <- pre_data %>%
  filter(experiment == "preference_pretest")%>%
  group_by(condition,age_num,subid)%>%
  summarise(correct = mean(correct))

ggplot(data = pre_prior, aes(x = age_num, y = correct, col = condition)) +
  geom_jitter(alpha = 0.5, height = 0.1, width = 0.1)+
  geom_smooth(aes(fill = condition),method = "lm", se = T, col = "black",  alpha = .5, size = 1)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="age",y="Proportion Expected Choice")+
  theme_few(base_size = 20) +
  #ggtitle("Preference data")+
  ylim(0,1)+
  guides(alpha = F)+ 
  coord_fixed(ratio = 2)+
  scale_color_solarized()+
  scale_fill_solarized()+
  theme(panel.border = element_rect(color = "#268bd2", fill = NA, size = 6),legend.position="bottom")

```{r}
ggsave("../graphs/kids_pref.png", width = 5, height = 4, scale = 1.5)
```

```{r pretest plot prior, echo = F}

kids_novel_data <- read_csv(file="../data/kids_novel.data.round1.csv")%>%
  filter(age != 2)%>%
  mutate(condition = ifelse(change == "false", "same_speaker", "different_speaker"))

p_kids_novel <- kids_novel_data %>%
  group_by(condition,age_num,subid)%>%
  summarise(correct = mean(correct))

ggplot(data = p_kids_novel, aes(x = age_num, y = correct, col = condition)) +
  geom_jitter(alpha = 0.5, height = 0.1, width = 0.1)+
  geom_smooth(aes(fill = condition),method = "lm", se = T, col = "black",  alpha = .5, size = 1)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="age",y="Proportion Expected Choice")+
  theme_few(base_size = 20) +
  #ggtitle("Preference data")+
  ylim(0,1)+
  guides(alpha = F)+ 
  coord_fixed(ratio = 2)+
  scale_color_solarized()+
  scale_fill_solarized()+
  theme(panel.border = element_rect(color = "#268bd2", fill = NA, size = 6),legend.position="bottom")

```{r}
ggsave("../graphs/kids_novel.png", width = 5, height = 4, scale = 1.5)
```

### Model based speaker optimality by age

```{r plot so model pred, echo = F}

plot_model_pred_so <- kids_model_pred_so%>%
  group_by(age_bin)%>%
  summarise(mean = mean(speakerOptimality),
            lci = hdi_lower(speakerOptimality),
            uci = hdi_upper(speakerOptimality))

ggplot(data = plot_model_pred_so) +
  geom_line(data = kids_model_pred_so,aes(age_bin, speakerOptimality, group = iteration), col = "grey", alpha = 0.3) +
  geom_ribbon(aes(x =age_bin, ymin = lci, ymax = uci), fill = "green", alpha = .2) +
  geom_line(aes(age_bin, mean), col = "black", size = 1, alpha = 1) + 
  scale_x_continuous(limits=c(0,2), name="age", breaks = c(0,0.5,1,1.5,2), labels = c(3.0,3.5,4.0,4.5,5)) +
  scale_y_continuous(limits=c(-4, 4), name="Speaker optimality")+
  geom_hline(yintercept = 0, lty=2)+
  labs(x="age",y="Proportion Expected Choice")+
  theme_few(base_size = 20) +
  #ggtitle("Informativeness data")+
  ylim(-4,4)+
  guides(alpha = F)+ 
  coord_fixed(ratio = 1/4)+
  theme(panel.border = element_rect(color = "#dc322f", fill = NA, size = 6))

```{r}
ggsave("../graphs/kids_so_model.png", width = 3, height = 3, scale = 1.5)
```

```{r plot prior model pred, echo = F}
plot_model_pred_prior <- kids_model_pred_prior%>%
  group_by(age_bin,condition)%>%
  summarise(mean = mean(prop_preferred),
            lci = hdi_lower(prop_preferred),
            uci = hdi_upper(prop_preferred))

ggplot(data = p_kids_novel, aes(x = age_num, y = correct, col = condition)) +
  geom_jitter(alpha = 0.5, height = 0.1, width = 0.1)+
  geom_smooth(aes(fill = condition),method = "lm", se = T, col = "black",  alpha = .5, size = 1)+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="age",y="Proportion Expected Choice")+
  theme_few(base_size = 20) +
  #ggtitle("Preference data")+
  ylim(0,1)+
  guides(alpha = F)+ 
  coord_fixed(ratio = 2)+
  scale_color_solarized()+
  scale_fill_solarized()+
  theme(panel.border = element_rect(color = "#268bd2", fill = NA, size = 6),legend.position="bottom")

ggplot(data = plot_model_pred_prior) +
  geom_line(data = kids_model_pred_prior,aes(age_bin, prop_preferred, group = paste(iteration,condition, sep= ""), col = condition), alpha = 0.05) +
  geom_ribbon(aes(x =age_bin, ymin = lci, ymax = uci, fill = condition), alpha = .5) +
  geom_line(aes(age_bin, mean, group = condition), col = "black", size = 1, alpha = 1) + 
  scale_x_continuous(limits=c(0,2), name="age", breaks = c(0,0.5,1,1.5,2), labels = c(3.0,3.5,4.0,4.5,5)) +
  scale_y_continuous(limits=c(-0.05,1.05), name="Proportion Expected Choice")+
  geom_hline(yintercept = 0.5, lty=2)+
  labs(x="age",y="Proportion Expected Choice")+
  theme_few(base_size = 20) +
  guides(alpha = F)+ 
  coord_fixed(ratio = 2)+
  scale_color_solarized()+
  scale_fill_solarized()+
  theme(panel.border = element_rect(color = "#268bd2", fill = NA, size = 6),legend.position="bottom")

```{r}
ggsave("../graphs/kids_prior_model.png", width = 5, height = 4, scale = 1.5)
```

```{r}
model_comp_mean_ci <- bind_rows(
#plot_model_pred%>%ungroup()%>%mutate(model = "RSA"), 
plot_prior_only_pred%>%ungroup()%>%mutate(model = "Prior Only"),
no_prior_plot%>%ungroup()%>%mutate(model = "No Prior")
)

model_comp <- bind_rows(
#kids_model_pred %>%mutate(model = "RSA"),
prior_only_model%>%mutate(model = "Prior Only"),
no_prior_pred%>%select(-so_slope,-so_int,-speakerOptimality)%>%mutate(model = "No Prior")
  )

```{r, include = F}
ggplot(model_comp_mean_ci)+
  geom_line(data = model_comp,aes(age_bin, prop_informative, group = paste(iteration,Speaker,model, sep= ""), col = model),  alpha = 0.01) + 
  geom_ribbon(aes(x =age_bin, ymin = lci, ymax = uci, fill = model), alpha = .4) +
  geom_line(aes(age_bin, mean, col = model), size = 1, alpha = 1) + 
  geom_hline(yintercept = 0.5, lty=2)+
  labs(y="proportion more informative")+
  scale_x_continuous(limits=c(0,2), name="age", breaks = c(0,0.5,1,1.5,2), labels = c(3.0,3.5,4.0,4.5,5)) +
  ylim(-0.05,1.05)+
  facet_grid(Speaker~Alignment)+
  theme_few(base_size = 20)+
  scale_color_manual( values = c("#dc322f","#268bd2"))+
  scale_fill_manual(values = c("#dc322f","#268bd2"))
```

```{r, include = F}
ggsave("../graphs/kids_no_prior_prior_only_model_predictions.pdf", width = 9, height = 7, scale = 1.2)
```
